---
output:
  pdf_document: default
  html_document: default
---
\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
getwd()
# Load your packages
library(tidyverse)
library(plyr)
library(lubridate)
library(trend)
library(here)
library(openmeteo)
# Set your ggplot theme

# Load your datasets
Stanislaus_Air_Files <- list.files(path = "./Data/Raw/Stanislaus_EPAair/",
                                  pattern = "*.csv",
                                  full.names = TRUE)

Stanislaus_Air_Data <- Stanislaus_Air_Files %>% 
  plyr::ldply(read.csv)


#Load weather data

install.packages("openmeteo") 
library(openmeteo)

weather_data <- weather_history(
 location = c(37.6393,120.9970), #Coordinates for Modesto, CA
 start = "2013-01-01", 
 end = "2023-12-31",
 hourly = "temperature_2m" 
)


```

```{r weatherdataprocess}
# Ensure plyr is detached to avoid conflicts
if ("package:plyr" %in% search()) {
  detach("package:plyr", unload = TRUE)
}

# Ensure datetime is in proper format
weather_data <- weather_data %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S"))

# Combine rows by month
weather_data_processed <- weather_data %>%
  mutate(
    month_year = format(as.Date(datetime), "%Y-%m"), 
  ) %>%
  group_by(month_year) %>%
  summarise(
    avg_temperature = mean(hourly_temperature_2m, na.rm = TRUE), 
    max_temperature = max(hourly_temperature_2m, na.rm = TRUE), 
    min_temperature = min(hourly_temperature_2m, na.rm = TRUE), 
    total_records = n() 
  )
```
# Rationale and Research Questions




\newpage
```{r airqualityprocess}
# Process the data: Combine rows by month
Stanislaus_Air_Data_Processed <- Stanislaus_Air_Data %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%Y"), 
    month_year = format(Date, "%Y-%m")             
  ) %>%
  group_by(month_year) %>%                   
  summarise(
    avg_PM25 = mean(Daily.Mean.PM2.5.Concentration, na.rm = TRUE),  
    max_PM25 = max(Daily.Mean.PM2.5.Concentration, na.rm = TRUE), 
    min_PM25 = min(Daily.Mean.PM2.5.Concentration, na.rm = TRUE),  
    total_records = n()                                             
  )


Stanislaus_Air_Data_Processed

```

# Dataset Information
We downloaded two datasets, one from EPA Stanislaus_Air_Data to present the air 
quality(include two values we want to examine PM2.5 and AQI values), another is 
weather data from NOAA, we installed "openmeteo” package to process data. Since 
two datasets have different format of dates, we first use mutate function and group
_by month to combine the rows for ensuring the consistency of "yyyy%-mm%" format. 
However, each month has different numbers of observations, we then combine all the 
rows happen in the same month so that we can easier to analyze the data. 


\newpage

# Exploratory Analysis 

##Are there any noticeable trends in temperature or air pollution?

```{r}
#Explore data set
class(weather_data_processed$month_year)

#Create new column that is a date
weather_data_processed <- weather_data_processed %>%
  mutate(month_year_date = as.Date(paste0(month_year, "-01")))

Stanislaus_Air_Data_Processed<- Stanislaus_Air_Data_Processed %>%
  mutate(month_year_date = as.Date(paste0(month_year, "-01")))

#Verify 
str(weather_data_processed)
str(Stanislaus_Air_Data_Processed)


```

```{r}
#Create a joined data frame with air and temperature data 

combined_air_weather <- left_join(weather_data_processed, Stanislaus_Air_Data_Processed, by = "month_year_date")%>%
   slice(-1) #remove first column that contained NA values for Dec 2012

combined_data_long <- combined_air_weather %>%
  pivot_longer(cols = c(avg_temperature, avg_PM25), names_to = "variable", values_to = "value")

#Create a faceted plot of average temperature and average PM2.5
ggplot(combined_data_long, aes(x = month_year_date, y = value, color = variable)) +
  geom_line(size = 1) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+  # Adjust interval
  facet_wrap(~ variable, scales = "free_y", ncol = 1, 
            labeller = labeller(variable = c("avg_temperature" = "Average Temperature (°C)", 
                                              "avg_pm25" = "Average PM2.5 (µg/m³)"))) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    legend.position = "none"
  ) +
  labs(
    title = "Monthly Trends of Temperature and PM2.5")


```
###Commentary
Looking at this graph, it appears that in the last decade, temperatures in Stanislaus County have followed a regular seasonal pattern. It does not show any type of trend year over year. On the other hand, PM2.5 shows a large spike between 2018-2019 and some other smaller spikes which appear to occur at somewhat regular intervals. 

##How do PM2.5 concentrations vary over the course of a year? 

```{r}
#Create a new column with month and year
Stanislaus_Air_Data_Processed <- Stanislaus_Air_Data_Processed %>%
  mutate(
    year = format(as.Date(month_year_date), "%Y"),
    month = format(as.Date(month_year_date), "%m")
  )

#Plot PM2.5 levels by year

ggplot(Stanislaus_Air_Data_Processed, 
       aes(x = as.numeric(month), 
           y = avg_PM25, 
           color = year, 
           group = year)) +
  geom_line() +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Show months as x-axis labels
  labs(
    title = "Monthly PM2.5 Concentrations, From Years 2013-2023",
    x = "Month",
    y = "PM2.5 Concentrations (µg/m³)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), 
    legend.title = element_blank() 
  )
```

###Commentary

It looks like from 2013-2023, there is generally an increase in PM2.5 from July onwards, then it dips back down during teh winter months. 


It is possible that there could be a seasonal component to particulate matter. We know that anthropogenic sources of PM 2.5 include combustion from motor vehicles, smelters, power plants, industrial facilities, residential fireplaces/wood stoves, agricultural burning, and forest fires. Thus, it is possible that in California, there could be higher PM 2.5 in the summer, when the risk of forest fires is higher and power plants are burning more fossil fuels to meet peak demand for air conditioning during hot summer months. 


##
\newpage

# Analysis

## Question 1: Do concentrations of PM 2.5 correlate with temperature?

## Question 2: 
\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
