---
output:
  pdf_document: default
  html_document: default
---
\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
getwd()
# Load your packages
library(tidyverse)
library(plyr)
library(lubridate)
library(trend)
library(here)
library(openmeteo)
# Set your ggplot theme

# Load your datasets
Stanislaus_Air_Files <- list.files(path = "./Data/Raw/Stanislaus_EPAair/",
                                  pattern = "*.csv",
                                  full.names = TRUE)

Stanislaus_Air_Data <- Stanislaus_Air_Files %>% 
  plyr::ldply(read.csv)
<<<<<<< HEAD
 
weather_data <- weather_history(
 location = c(36.000,-78.900),
 start = "2020-01-01",
 end = "2021-12-31",
 hourly = "temperature_2m"
=======

#Load weather data

install.packages("openmeteo") 
library(openmeteo)

weather_data <- weather_history(
 location = c(36.000,-78.900), #change these coordinates
 start = "2020-01-01", #change the dates
 end = "2021-12-31",
 hourly = "temperature_2m" 
>>>>>>> refs/remotes/origin/main
)
 
ggplot(weather_data,aes(x=datetime,y=hourly_temperature_2m)) + 
  geom_line()
<<<<<<< HEAD
=======


>>>>>>> refs/remotes/origin/main
```

```{r weatherdataprocess}
# Ensure plyr is detached to avoid conflicts
if ("package:plyr" %in% search()) {
  detach("package:plyr", unload = TRUE)
}

# Ensure datetime is in proper format
weather_data <- weather_data %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S"))

# Combine rows by month
weather_data_processed <- weather_data %>%
  mutate(month = format(as.Date(datetime), "%Y-%m")) %>% # Extract month as "YYYY-MM"
  group_by(month) %>% 
  summarise(
    avg_temperature = mean(hourly_temperature_2m, na.rm = TRUE), 
    max_temperature = max(hourly_temperature_2m, na.rm = TRUE), 
    min_temperature = min(hourly_temperature_2m, na.rm = TRUE), 
    total_records = dplyr::n() # Explicitly use n() from dplyr
  )

weather_data_processed
```
# Rationale and Research Questions




\newpage
```{r airqualityprocess}
# Process the data: Combine rows by month
Stanislaus_Air_Data_Processed <- Stanislaus_Air_Data %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%Y"), 
    month = format(Date, "%Y-%m")             
  ) %>%
  group_by(month) %>%                   
  summarise(
    avg_PM2_5 = mean(Daily.Mean.PM2.5.Concentration, na.rm = TRUE),  
    max_PM2_5 = max(Daily.Mean.PM2.5.Concentration, na.rm = TRUE), 
    min_PM2_5 = min(Daily.Mean.PM2.5.Concentration, na.rm = TRUE),  
    total_records = n()                                             
  )


Stanislaus_Air_Data_Processed

```

# Dataset Information
We are download two dataset, one from EPA Stanislaur_Air_Data to present the air 
quality(include two values we want to examine PM2.5 and AQI values), another is 
weather data from NOAA, we installed openmeteo” package to process data. Since 
two datasets have different format of dates, we first use mutate function and group
_by month to combine the rows for ensuring the consistency of "yyyy%-mm%" format. 
However, each month has differnt numbers of observations, we then combine all the 
rows happen in the same month so that we can easier to analyze the data. 

weather_data to see how 


\newpage

# Exploratory Analysis 



\newpage

# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
